Non-word repetition in children learning Yélî Dnye

### 2021-10-18 manuscript.Rmd
add missing refs + re-knit

lang              : "english"
class             : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    includes:
      in_header: header.tex
---

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links  between phonology, working memory, and the lexicon are [@bowey2001nonword], and how extensively phonological constraints found in the lexicon affect online production [@gallagher2014identity]. NWR is also frequently used in applied contexts, notably as a diagnostic tool for language delays and disorders [@chiat2015non;@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe in the context of diagnosing language impairments among bilingual children [@meir2016disentangling;@is08042009language;@chiat2015nonword;@armon2015assessing]. NWR tasks probably tap into many skills [for relevant discussion see @coady2008uses; @dos2020epreuve]. Non-words can be designed to try to isolate certain skills more narrowly; for instance, one can choose non-words that contain real morphemes in order to load more on prior language experience, or non-words that are shorter to avoid loading on working memory [see a discussion in @chiat2015nonword]. Broadly, however, NWR scores will necessarily reflect to a certain extent phonological knowledge (to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

@@

The first research area is at the intersection of typology and phonological development. There has been an interest in adapting NWR to different languages, in part for applied purposes. In a review of NWR as a potential task to diagnose language impairments among bilingual children in Europe, @chiat2015non discusses the impossibility of creating language-universal non-word items: Languages vary in their phonological inventory, sound sequencing (phonotactics), syllable structure, and word-level prosody. As a result, any one item created will be relatively easier if it more closely resembles real words in a language, making it difficult to balance difficulty when comparing children learning different languages. This previous literature also suggests some dimensions of difficulty—an issue to which we return in the next subsection.

Although this cross-linguistic literature is rich, the potential difficulty associated with specific phonetic targets composing the non-words has received relatively little attention. For example, @chiat2015non discusses segmental complexity as a function of whether there are consonant clusters -- which is arguably a factor reflecting phonotactics and syllable structure.

@@

It is possible that differences are due to language-specific characteristics, including the most common length of words in the lexicon and/or in child-experienced speech in that culture—a hypothesis discussed for instance in @chiat2015non (pp. 7-8; see also p. 5). In broad terms, one may expect languages with a lexicon that is heavily biased towards monosyllables to show greater length effects than languages where words tend to be longer. A non-systematic meta-analysis does not provide overwhelming support for this hypothesis [@cristia2021supplementary; SM1].

@@

<!-- ac2mc shouldn't we totally rewrite this? // actually, this is all factually correct-->
<!-- mc2ac: while factually correct i thought it might not be a great representation of where we stand now so i made some small edits in the final sentences-->
We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech [@cristia2020infant]. However, @cristia2020infant also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is comparably infrequent in this community [in fact it may be infrequent in many settings, including urban ones @bunceURcross], and additionally shares other societal characteristics wih the Tsimane' [e.g., is rural and relies on farming, children grow up in wide familial networks, etc; @casillas2020early]. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour [@scaff2021daylong], and Yélî children about 3.6 minutes per hour [@casillas2020early]. Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in @cristia2020infant.

>>>>

lang              : "en-US"
class             : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    includes:
      in_header: header.tex
---

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links  between phonology, working memory, and the lexicon are [@bowey2001nonword], and how extensively phonological constraints found in the lexicon affect online production [@gallagher2014identity]. NWR is also frequently used in applied contexts, notably as a diagnostic tool for language delays and disorders [@chiat2015nonword;@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe in the context of diagnosing language impairments among bilingual children [@meir2016disentangling;@is08042009language;@chiat2015nonword;@armon2015assessing]. NWR tasks probably tap into many skills [for relevant discussion see @coady2008uses; @dos2020epreuve]. Non-words can be designed to try to isolate certain skills more narrowly; for instance, one can choose non-words that contain real morphemes in order to load more on prior language experience, or non-words that are shorter to avoid loading on working memory [see a discussion in @chiat2015nonword]. Broadly, however, NWR scores will necessarily reflect to a certain extent phonological knowledge (to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

@@

The first research area is at the intersection of typology and phonological development. There has been an interest in adapting NWR to different languages, in part for applied purposes. In a review of NWR as a potential task to diagnose language impairments among bilingual children in Europe, @chiat2015nonword discusses the impossibility of creating language-universal non-word items: Languages vary in their phonological inventory, sound sequencing (phonotactics), syllable structure, and word-level prosody. As a result, any one item created will be relatively easier if it more closely resembles real words in a language, making it difficult to balance difficulty when comparing children learning different languages. This previous literature also suggests some dimensions of difficulty—an issue to which we return in the next subsection.

Although this cross-linguistic literature is rich, the potential difficulty associated with specific phonetic targets composing the non-words has received relatively little attention. For example, @chiat2015nonword discusses segmental complexity as a function of whether there are consonant clusters -- which is arguably a factor reflecting phonotactics and syllable structure.

@@

It is possible that differences are due to language-specific characteristics, including the most common length of words in the lexicon and/or in child-experienced speech in that culture—a hypothesis discussed for instance in @chiat2015nonword (pp. 7-8; see also p. 5). In broad terms, one may expect languages with a lexicon that is heavily biased towards monosyllables to show greater length effects than languages where words tend to be longer. A non-systematic meta-analysis does not provide overwhelming support for this hypothesis [@cristia2021supplementary; SM1].

@@

<!-- ac2mc shouldn't we totally rewrite this? // actually, this is all factually correct-->
<!-- mc2ac: while factually correct i thought it might not be a great representation of where we stand now so i made some small edits in the final sentences-->
We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech [@cristia2020infant]. However, @cristia2020infant also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is comparably infrequent in this community [in fact it may be infrequent in many settings, including urban ones @bunceURcrosscultural], and additionally shares other societal characteristics wih the Tsimane' [e.g., is rural and relies on farming, children grow up in wide familial networks, etc; @casillas2020early]. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour [@scaff2021daylong], and Yélî children about 3.6 minutes per hour [@casillas2020early]. Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in @cristia2020infant.

### 2021-10-18 manuscript.tex
add missing refs + re-knit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  ,man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Non-word repetition in children learning Yélî Dnye},
  pdfauthor={Alejandrina Cristia1 \& Marisa Casillas2,3},
  pdflang={en-EN},
  pdfkeywords={phonology, non-word repetition, Papuan, non-industrial, non-urban, comparative, typology, markedness, literacy},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{fontspec}

@@

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and how extensively phonological constraints found in the lexicon affect online production (Gallagher, 2014). NWR is also frequently used in applied contexts, notably as a diagnostic tool for language delays and disorders (Estes, Evans, \& Else-Quest, 2007; \textbf{chiat2015non?}). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe in the context of diagnosing language impairments among bilingual children (Armon-Lotem, Jong, \& Meir, 2015; Chiat, 2015; COST Action, 2009; Meir, Walters, \& Armon-Lotem, 2016). NWR tasks probably tap into many skills (for relevant discussion see Coady \& Evans, 2008; Santos, Frau, Labrevoit, \& Zebib, 2020). Non-words can be designed to try to isolate certain skills more narrowly; for instance, one can choose non-words that contain real morphemes in order to load more on prior language experience, or non-words that are shorter to avoid loading on working memory (see a discussion in Chiat, 2015). Broadly, however, NWR scores will necessarily reflect to a certain extent phonological knowledge (to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

@@

The first research area is at the intersection of typology and phonological development. There has been an interest in adapting NWR to different languages, in part for applied purposes. In a review of NWR as a potential task to diagnose language impairments among bilingual children in Europe, (\textbf{chiat2015non?}) discusses the impossibility of creating language-universal non-word items: Languages vary in their phonological inventory, sound sequencing (phonotactics), syllable structure, and word-level prosody. As a result, any one item created will be relatively easier if it more closely resembles real words in a language, making it difficult to balance difficulty when comparing children learning different languages. This previous literature also suggests some dimensions of difficulty---an issue to which we return in the next subsection.

Although this cross-linguistic literature is rich, the potential difficulty associated with specific phonetic targets composing the non-words has received relatively little attention. For example, (\textbf{chiat2015non?}) discusses segmental complexity as a function of whether there are consonant clusters -- which is arguably a factor reflecting phonotactics and syllable structure.

In the present study, we thought it was relevant to represent the rich phonological inventory found in Yélî Dnye, by including a variety of phonetic targets. Some of them are cross-linguistically rare, in that they are less common across languages than other sounds or phonetic targets. Phonologists, phoneticians, and psycholinguists have discussed the extent to which cross-linguistic frequency may reflect ease of processing and acquisition via diachronic language change. These works focus largely on phonotactics (Moreton \& Pater, 2012) perceptual parsing of the (ambiguous) linguistic signal (\textbf{beddor2009coarticulatory?}; \textbf{ohala1981listener?}), and individual differences in processing styles (\textbf{bermudezotero2015amphichronic?}); small but significant effects that may cumulatively drive language change via phonologization (see \textbf{yu2021toward?} for a recent review). Thus, the correlation between typological frequency and ease of acquisition is typically assumed to emerge from one or more of the following causal paths:

@@

It is possible that differences are due to language-specific characteristics, including the most common length of words in the lexicon and/or in child-experienced speech in that culture---a hypothesis discussed for instance in (\textbf{chiat2015non?}) (pp.~7-8; see also p.~5). In broad terms, one may expect languages with a lexicon that is heavily biased towards monosyllables to show greater length effects than languages where words tend to be longer. A non-systematic meta-analysis does not provide overwhelming support for this hypothesis {[}Cristia and Casillas (2021); SM1{]}.

@@

We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech (Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020). However, Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020) also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is comparably infrequent in this community (in fact it may be infrequent in many settings, including urban ones \textbf{bunceURcross?}), and additionally shares other societal characteristics wih the Tsimane' {[}e.g., is rural and relies on farming, children grow up in wide familial networks, etc; Casillas, Brown, and Levinson (2020){]}. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour (Scaff, Stieglitz, Casillas, \& Cristia, 2021), and Yélî children about 3.6 minutes per hour (Casillas, Brown, \& Levinson, 2020). Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020).

@@

We thus at present conclude that typological frequency of sounds is, to a certain extent, mirrored in children's NWR, in ways that may not be due merely to how often those sounds are used in the ambient language, and which are not erased by language-specific pressure to make finer-grained differences early in development. We do not aim to reopen a debate on the extent to which cross-linguistic frequency of occurrence can be viewed necessarily as reflecting ease of perception or production (Maddieson, 2009; via phonotactic constraints, ambiguous parsing conditions, individual differences, and more as in, e.g., \textbf{beddor2009coarticulatory?}; \textbf{bermudezotero2015amphichronic?}; \textbf{ohala1981listener?}; \textbf{yu2021toward?}), but we do point out that this association is interestingly different from effects found in artificial language learning tasks (see Moreton \& Pater, 2012 for a review) which are in some ways quite similar to NWR. We believe that it may be insightful to extend the purview of NWR from a narrow focus on working memory and structural factors to broader uses, including for describing the phonological representations in the perception-production loop (as in e.g., Edwards, Beckman, \& Munson, 2004).

@@

@@

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  american,
  ,man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Non-word repetition in children learning Yélî Dnye},
  pdfauthor={Alejandrina Cristia1 \& Marisa Casillas2,3},
  pdflang={en-US},
  pdfkeywords={phonology, non-word repetition, Papuan, non-industrial, non-urban, comparative, typology, markedness, literacy},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{fontspec}

@@

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\else
  \usepackage[main=american]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and how extensively phonological constraints found in the lexicon affect online production (Gallagher, 2014). NWR is also frequently used in applied contexts, notably as a diagnostic tool for language delays and disorders (Chiat, 2015; Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe in the context of diagnosing language impairments among bilingual children (Armon-Lotem, Jong, \& Meir, 2015; Chiat, 2015; COST Action, 2009; Meir, Walters, \& Armon-Lotem, 2016). NWR tasks probably tap into many skills (for relevant discussion see Coady \& Evans, 2008; Santos, Frau, Labrevoit, \& Zebib, 2020). Non-words can be designed to try to isolate certain skills more narrowly; for instance, one can choose non-words that contain real morphemes in order to load more on prior language experience, or non-words that are shorter to avoid loading on working memory (see a discussion in Chiat, 2015). Broadly, however, NWR scores will necessarily reflect to a certain extent phonological knowledge (to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

@@

The first research area is at the intersection of typology and phonological development. There has been an interest in adapting NWR to different languages, in part for applied purposes. In a review of NWR as a potential task to diagnose language impairments among bilingual children in Europe, Chiat (2015) discusses the impossibility of creating language-universal non-word items: Languages vary in their phonological inventory, sound sequencing (phonotactics), syllable structure, and word-level prosody. As a result, any one item created will be relatively easier if it more closely resembles real words in a language, making it difficult to balance difficulty when comparing children learning different languages. This previous literature also suggests some dimensions of difficulty---an issue to which we return in the next subsection.

Although this cross-linguistic literature is rich, the potential difficulty associated with specific phonetic targets composing the non-words has received relatively little attention. For example, Chiat (2015) discusses segmental complexity as a function of whether there are consonant clusters -- which is arguably a factor reflecting phonotactics and syllable structure.

In the present study, we thought it was relevant to represent the rich phonological inventory found in Yélî Dnye, by including a variety of phonetic targets. Some of them are cross-linguistically rare, in that they are less common across languages than other sounds or phonetic targets. Phonologists, phoneticians, and psycholinguists have discussed the extent to which cross-linguistic frequency may reflect ease of processing and acquisition via diachronic language change. These works focus largely on phonotactics (Moreton \& Pater, 2012) perceptual parsing of the (ambiguous) linguistic signal (Beddor, 2009; Ohala, 1981), and individual differences in processing styles (Bermúdez-Otero, 2015); small but significant effects that may cumulatively drive language change via phonologization (see Yu, 2021 for a recent review). Thus, the correlation between typological frequency and ease of acquisition is typically assumed to emerge from one or more of the following causal paths:

@@

It is possible that differences are due to language-specific characteristics, including the most common length of words in the lexicon and/or in child-experienced speech in that culture---a hypothesis discussed for instance in Chiat (2015) (pp.~7-8; see also p.~5). In broad terms, one may expect languages with a lexicon that is heavily biased towards monosyllables to show greater length effects than languages where words tend to be longer. A non-systematic meta-analysis does not provide overwhelming support for this hypothesis {[}Cristia and Casillas (2021); SM1{]}.

@@

We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech (Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020). However, Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020) also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is comparably infrequent in this community (in fact it may be infrequent in many settings, including urban ones Bunce et al., under review), and additionally shares other societal characteristics wih the Tsimane' {[}e.g., is rural and relies on farming, children grow up in wide familial networks, etc; Casillas, Brown, and Levinson (2020){]}. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour (Scaff, Stieglitz, Casillas, \& Cristia, 2021), and Yélî children about 3.6 minutes per hour (Casillas, Brown, \& Levinson, 2020). Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020).

@@

We thus at present conclude that typological frequency of sounds is, to a certain extent, mirrored in children's NWR, in ways that may not be due merely to how often those sounds are used in the ambient language, and which are not erased by language-specific pressure to make finer-grained differences early in development. We do not aim to reopen a debate on the extent to which cross-linguistic frequency of occurrence can be viewed necessarily as reflecting ease of perception or production (via phonotactic constraints, ambiguous parsing conditions, individual differences, and more as in, e.g., Beddor, 2009; Bermúdez-Otero, 2015; Maddieson, 2009; Ohala, 1981; Yu, 2021), but we do point out that this association is interestingly different from effects found in artificial language learning tasks (see Moreton \& Pater, 2012 for a review) which are in some ways quite similar to NWR. We believe that it may be insightful to extend the purview of NWR from a narrow focus on working memory and structural factors to broader uses, including for describing the phonological representations in the perception-production loop (as in e.g., Edwards, Beckman, \& Munson, 2004).

@@

\leavevmode\hypertarget{ref-beddor2009coarticulatory}{}%
Beddor, P. S. (2009). A coarticulatory path to sound change. \emph{Language}, \emph{85}(4), 785--832.

\leavevmode\hypertarget{ref-bermudezotero2015amphichronic}{}%
Bermúdez-Otero, R. (2015). Amphichronic explanation and the life cycle of phonological processes. In P. Honeybone \& J. Salmons (Eds.), \emph{The {O}xford handbook of historical phonology} (pp. 374--399). Oxford, UK: Oxford University Press.

@@

\leavevmode\hypertarget{ref-bunceURcrosscultural}{}%
Bunce, J., Soderstrom, M., Bergelson, E., Rosemberg, C., Stein, A., Alam, F., \ldots{} Casillas, M. (under review). A cross-cultural examination of young children's everyday language experiences.

@@

\leavevmode\hypertarget{ref-ohala1981listener}{}%
Ohala, J. J. (1981). The listener as a source of sound change. In M. F. Miller, C. S. Masek, \& R. A. Hendrick (Eds.), \emph{Papers from the parasession on language and behavior} (pp. 178--203). Chicago, IL: Chicago Linguistics Society.

@@

\leavevmode\hypertarget{ref-yu2021toward}{}%
Yu, A. C. L. (2021). Toward an individual-difference perspective on phonologization. \emph{Glossa: A Journal of General Linguistics}, \emph{6}(1), 1--24.
